Mg Depleted and K Enriched Stars from LAMOST Spectral Data

### 2018-01-10 article/ms.tex
reordering; temp comments

@@

While it is plausible that the increased mass and greater distance from the gaseous disk of the Milky Way would cause a clusters first generation polluter stars to have an increased effect on the second generation of stars, this effect would apply to all ejecta (assuming comparable energies). So while it might be expected that second generation stars in clusters similar to NGC 2419 made up of more material processed by the first generation than other lighter globular clusters, it doesn't explain why the ejecta from the first generation was apparently dominated by this as yet unknown polluter object resulting in the Mg-K anti-correlation. In the opinion of the author, it is the identity and physical characterisation of this unknown polluter that is of the greatest scientific interest, with such a result also hopefully answering the question of the signatures uniqueness.

An early modelling attempt at replicating the anomalous abundances in NGC 2419 assumed hot-bottom burning in AGB and super-AGB (SAGB) stars \citep{ventura2012}. This study succeeded in reproducing the Mg-K anti-correlation, using the nuclear reaction pathway \ce{^{36}Ar(p,\gamma)^{37}K(\beta ^+ \nu)^{37}Ar(e^-,\nu)^{37}Cl(p,\gamma)^{38}Ar(p,\gamma)^39K} (corrected as per \cite{iliadis2016}) for nuclear synthesis of K. However, abundance patterns consistent with NGC 2419 were only attainable if the reaction cross section was set to 100 times the commonly accepted rate or temperatures could be made higher than currently accepted values, achieving a burning temperature of 150 MK.

A more recent, comprehensive attempt to replicate the Mg and K abundances in NGC 2419 by \cite{iliadis2016} also attempted to reproduce the abundances of Si, Sc, Ca, Ti and V (elements reported as having weak correlations with Mg by \cite{cohenkirby2012}). The ultimate goal was to constrain the temperatures and densities required to produce the chemical signiature, and thereby provide insight into potential polluter candidates. It is perhaps noteworthy that they identified a diferent main reaction pathway for K nucleosynthesis to that of \cite{ventura2012}, which is relegated to a minor minor seccondary pathway. The main pathway they identify is \ce{^{36}Ar(p,\gamma)^{37}K(\beta ^+ \nu)^{37}Ar(p,\gamma)^{38}K(\beta ^+ \nu)^{38}Ar(p,\gamma)^39K}.

It was found that (with the exception of extremely high densities > $10^8$ g/cm$^3$) temperatures between 100 and 200 MK and densities between 10$^{-4}$ and 10$^8$ g/cm$^3$ were required. Using this information, core and shell burning of low mass, high mass and super massive stars were ruled out as potential poluter candidates, as well as regular AGB stars. Super-AGB (SAGB) stars however were regarded as potential candidates, with only a relatively small (roughly 10-20 MK) increase in temperature required to fall in the acceptable band of parameter space identified. The other potential candidate identified was Novae, although the lack of detailed models of white dwarfs accreting material similar to NGC 2419 either from a binary companion or the intra cluster medium limit appraisals of its feasibility regarding whether enough ejecta could be produced and retained by the cluster. However, based on current novae frequency in globular clusters determined by \cite{kato2013novae}, \cite{iliadis2016} conclude that the amount of material that would be produced by Novae is at most 1 \% of the total required mass to pollute 30 \% of NGC 2419.

Another proposed  polluter candidate is a pair instability super nova (PISN) \citep{carretta2013}. Unique to extremely massive population III stars, these events involve the total destruction of the star, with no black hole remnant being left behind. The main argument for this idea is that the extreme rarity of these events, coupled with the huge masses of processed material involved, could potentially allow the entire signature in NGC 2419 to be the result of a single event, explaining why it is not seen in any other globular clusters. Further, the signature odd-even proton number abundance pattern associated with PISNs could perhaps explain the anti-correlation between Mg and K, as well as the enhancements in Sc and V observed by \cite{cohenkirby2012}.

This study, which queries 454180 LAMOST giants for depleted Mg and enhanced K, seeks to provide additional information which can be used to guide and test models attempting to explain the Mg-K anti-correlation and related phenomena. The large sample size provides greater insight into the supposed exclusivity of the signature, and the stars identified present an important opportunity for follow up observations.

>>>>

\newcommand{\LamostGiants}{454180}
\newcommand{\project}[1]{\emph{#1}}
\newcommand{\lamost}{\project{LAMOST}

\newcommand{\teff}{T_{\rm eff}}
\newcommand{\logg}{\log_{10}[g\,({\rm cm\,s}^{-2})]}

@@

%While it is plausible that the increased mass and greater distance from the gaseous disk of the Milky Way would cause a clusters first generation polluter stars to have an increased effect on the second generation of stars, this effect would apply to all ejecta (assuming comparable energies). So while it might be expected that second generation stars in clusters similar to NGC 2419 made up of more material processed by the first generation than other lighter globular clusters, it doesn't explain why the ejecta from the first generation was apparently dominated by this as yet unknown polluter object resulting in the Mg-K anti-correlation. In the opinion of the author, it is the identity and physical characterisation of this unknown polluter that is of the greatest scientific interest, with such a result also hopefully answering the question of the signatures uniqueness.

%An early modelling attempt at replicating the anomalous abundances in NGC 2419 assumed hot-bottom burning in AGB and super-AGB (SAGB) stars \citep{ventura2012}. This study succeeded in reproducing the Mg-K anti-correlation, using the nuclear reaction pathway \ce{^{36}Ar(p,\gamma)^{37}K(\beta ^+ \nu)^{37}Ar(e^-,\nu)^{37}Cl(p,\gamma)^{38}Ar(p,\gamma)^39K} (corrected as per \cite{iliadis2016}) for nuclear synthesis of K. However, abundance patterns consistent with NGC 2419 were only attainable if the reaction cross section was set to 100 times the commonly accepted rate or temperatures could be made higher than currently accepted values, achieving a burning temperature of 150 MK.

%A more recent, comprehensive attempt to replicate the Mg and K abundances in NGC 2419 by \cite{iliadis2016} also attempted to reproduce the abundances of Si, Sc, Ca, Ti and V (elements reported as having weak correlations with Mg by \cite{cohenkirby2012}). The ultimate goal was to constrain the temperatures and densities required to produce the chemical signiature, and thereby provide insight into potential polluter candidates. It is perhaps noteworthy that they identified a diferent main reaction pathway for K nucleosynthesis to that of \cite{ventura2012}, which is relegated to a minor minor seccondary pathway. The main pathway they identify is \ce{^{36}Ar(p,\gamma)^{37}K(\beta ^+ \nu)^{37}Ar(p,\gamma)^{38}K(\beta ^+ \nu)^{38}Ar(p,\gamma)^39K}.

%It was found that (with the exception of extremely high densities > $10^8$ g/cm$^3$) temperatures between 100 and 200 MK and densities between 10$^{-4}$ and 10$^8$ g/cm$^3$ were required. Using this information, core and shell burning of low mass, high mass and super massive stars were ruled out as potential poluter candidates, as well as regular AGB stars. Super-AGB (SAGB) stars however were regarded as potential candidates, with only a relatively small (roughly 10-20 MK) increase in temperature required to fall in the acceptable band of parameter space identified. The other potential candidate identified was Novae, although the lack of detailed models of white dwarfs accreting material similar to NGC 2419 either from a binary companion or the intra cluster medium limit appraisals of its feasibility regarding whether enough ejecta could be produced and retained by the cluster. However, based on current novae frequency in globular clusters determined by \cite{kato2013novae}, \cite{iliadis2016} conclude that the amount of material that would be produced by Novae is at most 1 \% of the total required mass to pollute 30 \% of NGC 2419.

%Another proposed  polluter candidate is a pair instability super nova (PISN) \citep{carretta2013}. Unique to extremely massive population III stars, these events involve the total destruction of the star, with no black hole remnant being left behind. The main argument for this idea is that the extreme rarity of these events, coupled with the huge masses of processed material involved, could potentially allow the entire signature in NGC 2419 to be the result of a single event, explaining why it is not seen in any other globular clusters. Further, the signature odd-even proton number abundance pattern associated with PISNs could perhaps explain the anti-correlation between Mg and K, as well as the enhancements in Sc and V observed by \cite{cohenkirby2012}.

This study, which queries \LamostGiants\ \lamost\ giants for depleted Mg and enhanced K, seeks to provide additional information which can be used to guide and test models attempting to explain the Mg-K anti-correlation and related phenomena. The large sample size provides greater insight into the supposed exclusivity of the signature, and the stars identified present an important opportunity for follow up observations.
